---
title: "Interactive Network Visualization"
author: "Steven Surya Tanujaya"
date: "July 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here, after learn about tidygraph, ggraph, and community detection, the `visNetwork` package is introduced to create interactive network visualization. `visNetwork` is a R package for network visualization, using `vis.js` javascript library (http://visjs.org). All the remarks and bugs returns are welcome on github : https://github.com/datastorm-open/visNetwork.  

visNetwork works under `dplyr` nature and needs at least two informations :

a nodes *data.frame*, with `id` column
a edges *data.frame*, with `from` and `to` columns

Let us use the media dataset.

```{r}
library(tidygraph)
library(visNetwork)
media.edge<-read.csv("data/Dataset1-Media-Example-EDGES.csv")
media.node<-read.csv("data/Dataset1-Media-Example-NODES.csv")
visNetwork(media.node, media.edge, width = "100%")
```

visNetwork individual configuration is quite simple but needs certain specific column name in order to configure color, shape, size, etc. or some `vis*` function can also be used in order to accomodate global configuration. 

Node configuration (global)
```{r}
visNetwork(media.node, media.edge, width = "100%") %>% 
  visNodes(shape = "square", 
           color = list(background = "lightblue", 
                        border = "darkblue",
                        highlight = "yellow"),
           shadow = list(enabled = TRUE, size = 10)) %>%
  visLayout(randomSeed = 12) 
```

Node configuration (individual)

```{r}
nodes1 <- data.frame(media.node, 
                    shape = ifelse(media.node$type.label=="Newspaper","circle", "square"))
nodes1<-nodes1 %>% rename(label=media)

visNetwork(nodes1, media.edge, width = "100%") %>% 
  visNodes(color = list(background = "lightblue", 
                        border = "darkblue",
                        highlight = "yellow"),
           shadow = list(enabled = TRUE, size = 10))  %>%
  visLayout(randomSeed = 12)
```

Edge configuration (global)

```{r}
visNetwork(nodes1, media.edge, width = "100%") %>% 
  visEdges(shadow = TRUE,
           arrows =list(to = list(enabled = TRUE, scaleFactor = 2)),
           color = list(color = "lightblue", highlight = "red")) %>%
  visLayout(randomSeed = 12) # to have always the same network             
```

Edge configuration (individual)

```{r}
edges1<-media.edge %>% rename(label=type,width=weight) %>% mutate(color=ifelse(label=="hyperlink","yellow","blue"), arrows="to",width=width*0.3)

visNetwork(nodes = nodes1, edges = edges1, width = "100%")
```

Group of nodes and legend

```{r}
nodes2<-media.node %>% rename(group=type.label)
visNetwork(nodes2, edges1, width = "100%") %>% 
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend()
```

You can customize the shape using any icon or picture, click [here](http://datastorm-open.github.io/visNetwork/image_icon.html) for the detail.

Next, there is also `visOptions` that provides us with more interactive effect.

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  # darkblue square with shadow for group "A"
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  # red triangle for group "B"
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visOptions(highlightNearest = T)
```

or even MORE effect!!

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  # darkblue square with shadow for group "A"
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  # red triangle for group "B"
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visOptions(highlightNearest = list(enabled=T,degree=2,hover=T))
```

You can also highlight some nodes/edges based on your input.

Input by ID.

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  # darkblue square with shadow for group "A"
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  # red triangle for group "B"
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visOptions(highlightNearest = T,nodesIdSelection = T)
```

Input by Column.

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  # darkblue square with shadow for group "A"
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  # red triangle for group "B"
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visOptions(highlightNearest = T,selectedBy = "group")
```

We can also manipulating the network by yourself (in interactive way) just using `manipulate=T` without even type a single code!

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  # darkblue square with shadow for group "A"
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  # red triangle for group "B"
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visOptions(manipulation = T)
```

Last but not least, if we want to make it more interactive we can just add `visInteraction` and for more user friendly we can just add `visConfigure`.

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visInteraction(navigationButtons = T)
```
Reference:
- http://datastorm-open.github.io/visNetwork/
- https://cran.r-project.org/web/packages/visNetwork/vignettes/Introduction-to-visNetwork.html

________________________________________________________________
 
Additional: `visConfigure`

```{r}
visNetwork(nodes2, edges1, width = "100%") %>% 
  visGroups(groupname = "Newspaper", color = "lightblue", shape = "square", 
            shadow = list(enabled = TRUE)) %>% 
  visGroups(groupname = "TV", color = "red", shape = "triangle") %>% 
  visGroups(groupname = "Online", color = "green", shape = "circle") %>% 
  visLegend() %>% 
  visConfigure(enabled=T)
```
